---
title: "STAT697FINALEXAM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
setwd("C:/Users/flore/Documents/Fall 2019 - Spring 2020/STAT697/STAT697HW")

table = read.table("icu.dat", header = F)
dimnames(table)[[2]] <- c("ID", "STA", "AGE", "SEX", "RACE", "SER", "CAN", "CRN", "INF", "CPR", "SYS", "HRA", "PRE", "TYP", "FRA", "PO2", "PH", "PCO", "BIC", "CRE", "LOC")
table$RACE <- as.factor(table$RACE)
table$LOC <- as.factor(table$LOC)
```



## Question 1

Write down the equation for the logistic regression model of STA (Vital Status: 0 = Lived,1= Died) on AGE(Years).

```{r, echo=FALSE}
mod1 <- glm(STA ~ AGE, family=binomial, data=table)
summary(mod1)
```




The prediction equation  is:

$$
logit(\hat{\pi}) =  -3.05851 + 0.02754*AGE 
$$

$$
\pi(AGE) = \frac{exp(-3.05851 + 0.02754*AGE_i)}{1 + exp(-3.05851 + 0.02754*AGE_i)}
$$

Write down the log likelihood for the logistic regression model and the equation for solving the maximum likelihood estimates (MLE) of the parameters.


Log-likelihood for the logistic regression model:

$$
l(a,b) = \sum^{200}_{i=1} y_ilog(\pi) + (1-y_i)log(1-\pi) = \sum^{200}_{i=1} y_i log(\frac{\pi}{1-\pi}) + log(1-\pi)
$$

$$
log(\frac{\pi}{1-\pi}) = a + b*AGE_i
$$
$$ 
 1-\pi = 1- \frac{exp(a + b*AGE_i)}{1 + exp(a + b*AGE_i)} = (1 + exp(a + b*AGE_i))^{-1}
$$ 
 
$$
log(1-\pi) = log((1 + exp(a + b*AGE_i))^{-1})= -3.05851 + 0.02754*AGE_i
$$ 
$$ 
l(a, b)= \sum^{200}_{i=1} STA_i(a + b*AGE_i) - log(1+ exp(a + b*AGE_i))
$$ 
 
$$ 
l(a, b)= \sum^{200}_{i=1} STA_i(-3.05851 + 0.02754*AGE_i) - log(1+ exp(-3.05851 + 0.02754*AGE_i))
$$

Equation for solving the MLE of the parameters:

$$
\frac{\partial l(a, b)}{\partial a} = \sum^{200}_{i=1} (STA_i - \pi) = 0
$$
$$
\frac{\partial l(a, b)}{\partial b} = \sum^{200}_{i=1} AGE_i(STA_i-\pi) = 0
$$

## Question 2

We can observe the scatterplot of STA versus AGE

```{r, echo=FALSE}
library(ggplot2)
plot <- ggplot( data= table) + 
  geom_point(mapping = aes(x= AGE, y = STA)) +
      ggtitle('Scatter plot of the Vital status over Age') +
         theme(plot.title = element_text(hjust = 0.5))

plot
```



## Question 3

We compute the STA mean over subjects within each AGE interval:


```{r, include = FALSE}
n<-numeric(8)
STAsum <-numeric(8)

for(i in 1:200){
  if (table$AGE[i] < 25){
    n[1] <- n[1] +1
    STAsum[1] <- STAsum[1] +table$STA[i]
  }
  else if (24 < table$AGE[i] && table$AGE[i] < 35){
    n[2] <- n[2] +1
    STAsum[2] <- STAsum[2] +table$STA[i]
  }
  else if (34 < table$AGE[i] && table$AGE[i] < 45){
    n[3] <- n[3] +1
    STAsum[3] <- STAsum[3] +table$STA[i]
  }
  else if (44 < table$AGE[i] && table$AGE[i] < 55){
    n[4] <- n[4] +1
    STAsum[4] <- STAsum[4] +table$STA[i]
  }
  else if (54 < table$AGE[i] && table$AGE[i] < 65){
    n[5] <- n[5] +1
    STAsum[5] <- STAsum[5] +table$STA[i]
  }
  else if (64 < table$AGE[i] && table$AGE[i] < 75){
    n[6] <- n[6] +1
    STAsum[6] <- STAsum[6] +table$STA[i]
  }
  else if (74 < table$AGE[i] && table$AGE[i] < 85){
    n[7] <- n[7] +1
    STAsum[7] <- STAsum[7] +table$STA[i]
  }
  else if (table$AGE[i] > 84){
    n[8] <- n[8] +1
    STAsum[8] <- STAsum[8] +table$STA[i]
  }
  
}
n  # number of people per category
STA_mean = STAsum/n
STA_mean
```

The number of people per interval is:

26  8 11 25 39 50 30 11

and the percentage of people who died per interval is:

0.0769 0 0.1818 0.2 0.2051 0.18 0.3 0.4545

We plot these values of mean STA versus the midpoint of the AGE interval:

```{r, echo=FALSE}
Age<-c(19.5, 29.5, 39.5, 49.5, 59.5, 69.5, 79.5, 89.5)
library(ggplot2)
plot <- ggplot( data= table) + 
  geom_point(mapping = aes(x= AGE, y = STA, col = 'black')) +
      ggtitle('Vital Status over Age') +
         theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(aes(x = Age[1], y = STA_mean[1], col = 'green')) +
  geom_point(aes(x = Age[2], y = STA_mean[2], col = 'green')) +
  geom_point(aes(x = Age[3], y = STA_mean[3], col = 'green')) +
  geom_point(aes(x = Age[4], y = STA_mean[4], col = 'green')) +
  geom_point(aes(x = Age[5], y = STA_mean[5], col = 'green')) +
  geom_point(aes(x = Age[6], y = STA_mean[6], col = 'green')) +
  geom_point(aes(x = Age[7], y = STA_mean[7], col = 'green')) +
  geom_point(aes(x = Age[8], y = STA_mean[8], col = 'green'))
plot+
    scale_color_identity(breaks = c('black', 'green'),
                       labels = c('Response', 'Group Means'),
                       guide = 'legend')
```


## Question 4

We obtain the MLE of the parameters of mean STA versus the midpoint of the AGE interval.

recall:STA is the Vital Status: 0 = Lived,1= Died, AGE is in Years


```{r, echo=FALSE}
mod2 <- glm(STAsum/n ~ Age, family=binomial, data=table, weights = n)
summary(mod2)

```

The prediction equation  is:

$$
logit(\hat{\pi}) =  -3.08609 + 0.02779*Age
$$


```{r, echo=FALSE}
table$mod <- exp(-3.059 + 0.0275*table$AGE) / (1 + exp(-3.059 + 0.0275*table$AGE))
```

We plot the estimated probabilities of the STA mean over subjects within each AGE interval:

```{r, echo=FALSE}
plot + geom_point(aes(x=AGE, y= mod, col = 'red'), data = table) +
      scale_color_identity(breaks = c('black', 'green','red'),
                       labels = c('Response', 'Group Means', 
                                  'Estimated Probability'),
                       guide = 'legend')

```


Does it appear plausible that the logit of the survival probability is linear in AGE?

Yes, it visually appears plausible.


## Question 5


1st part

We compute the STA mean over subjects within each SYS interval:


```{r, include=FALSE}

Sys<- c(seq(from=min(table$SYS),to=max(table$SYS),by= (max(table$SYS) - min(table$SYS))/8 ))

n<-numeric(8)
STAsum <-numeric(8)

for(i in 1:200){
  if (Sys[1] <= table$SYS[i] && table$SYS[i] < Sys[2]){
    n[1] <- n[1] +1
    STAsum[1] <- STAsum[1] +table$STA[i]
  }
  if (Sys[2] <= table$SYS[i] && table$SYS[i] < Sys[3]){
    n[2] <- n[2] +1
    STAsum[2] <- STAsum[2] +table$STA[i]
  }
  if (Sys[3] <= table$SYS[i] && table$SYS[i] < Sys[4]){
    n[3] <- n[3] +1
    STAsum[3] <- STAsum[3] +table$STA[i]
  }
  if (Sys[4] <= table$SYS[i] && table$SYS[i] < trunc((Sys[4]+Sys[5])/2) ){
    n[4] <- n[4] +1
    STAsum[4] <- STAsum[4] +table$STA[i]
  }
  if (trunc((Sys[4]+Sys[5])/2) <= table$SYS[i] && table$SYS[i] < Sys[5]){
  n[5] <- n[5] +1
  STAsum[5] <- STAsum[5] +table$STA[i]
  }
  if (Sys[5] <= table$SYS[i] && table$SYS[i] < Sys[6]){
  n[6] <- n[6] +1
  STAsum[6] <- STAsum[6] +table$STA[i]
  }
  if (Sys[6] <= table$SYS[i] && table$SYS[i] < Sys[7]){
  n[7] <- n[7] +1
  STAsum[7] <- STAsum[7] +table$STA[i]
  }

  if (Sys[7] <= table$SYS[i]) {
    n[8] <- n[8] +1
    STAsum[8] <- STAsum[8] +table$STA[i]
  }
}
n  # number of people per category
STA_mean = STAsum/n
STA_mean
```

The number of people per interval is:

4 14 46 38 39 41 12  6

and the percentage of people who died per interval is:

0.75 0.6429 0.1087 0.2105 0.2051 0.1219 0.0833 0.1667



We obtain the MLE of the parameters of mean STA versus the midpoint of the SYS interval.




```{r, echo=FALSE}
mod <- glm(STA ~ SYS, family=binomial, data=table)
summary(mod)
```

The prediction equation  is:

$$
logit(\hat{\pi}) =  0.777195 + -0.017019*Sys
$$
We also transfrom the data taking the log:

```{r, echo=FALSE}
log_mod <-glm(STA ~ log(SYS), family=binomial, data=table)
summary(log_mod)
```


The prediction equation  is:

$$
logit(\hat{\pi}) =  9.3131 + -2.2250*Sys
$$

```{r, echo=FALSE}
table$mod <- exp(0.777195 -0.017019*table$SYS) / (1 + exp(0.777195 -0.017019*table$SYS))
table$log_mod <- exp(9.3131 -2.2250*log(table$SYS)) / (1 + exp(9.3131 -2.2250*log(table$SYS)))
```


We plot  scatterplot of STA versus SYS, the mean STA versus the midpoint of the SYS interval and the estimated probabilities.

```{r, echo=FALSE, warning=FALSE}
Sys <- c(36, 63.5, 91, 118.5, 132, 146 ,173.5, 201, 256)
Sys <- c( (Sys[1]+Sys[2])/2, (Sys[2]+Sys[3])/2,  (Sys[3]+Sys[4])/2, (Sys[4]+Sys[5])/2, (Sys[5]+Sys[6])/2, (Sys[6]+Sys[7])/2, (Sys[7]+Sys[8])/2, (Sys[8]+Sys[9])/2 )


plot <- ggplot(data= table) + 
  geom_point(mapping = aes(x= SYS, y = STA, col = 'black')) +
      ggtitle('Vital Status over Systolic Blood Pressure') +
         theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(aes(x = Sys[1], y = STA_mean[1], col = 'green')) +
  geom_point(aes(x = Sys[2], y = STA_mean[2], col = 'green')) +
  geom_point(aes(x = Sys[3], y = STA_mean[3], col = 'green')) +
  geom_point(aes(x = Sys[4], y = STA_mean[4], col = 'green')) +
  geom_point(aes(x = Sys[5], y = STA_mean[5], col = 'green')) +
  geom_point(aes(x = Sys[6], y = STA_mean[6], col = 'green')) +
  geom_point(aes(x = Sys[7], y = STA_mean[7], col = 'green')) +
  geom_point(aes(x = Sys[8], y = STA_mean[8], col = 'green'))

plot + geom_point(aes(x=SYS, y= mod, col = 'red'), data = table) +
  geom_point(aes(x=SYS, y= log_mod, col = 'orange'), data = table) +
      scale_color_identity(breaks = c('black', 'green','red', 'orange'),
                       labels = c('Response', 'Group Means', 
                                  'Estimated Probability', 'log transformation'),
                       guide = 'legend')
```

2nd part

We compute the STA mean over subjects within each HRA interval:

```{r, include=FALSE}

Hra<- c(seq(from=min(table$HRA),to=max(table$HRA),by= (max(table$HRA) - min(table$HRA))/8 ))

n<-numeric(7)
STAsum <-numeric(7)

for(i in 1:200){
  if (Hra[1] <= table$HRA[i] && table$HRA[i] < Hra[2]){
    n[1] <- n[1] +1
    STAsum[1] <- STAsum[1] +table$STA[i]
  }
  if (Hra[2] <= table$HRA[i] && table$HRA[i] < Hra[3]){
    n[2] <- n[2] +1
    STAsum[2] <- STAsum[2] +table$STA[i]
  }
  if (Hra[3] <= table$HRA[i] && table$HRA[i] < Hra[4]){
    n[3] <- n[3] +1
    STAsum[3] <- STAsum[3] +table$STA[i]
  }
  if (Hra[4] <= table$HRA[i] && table$HRA[i] < Hra[5]){
    n[4] <- n[4] +1
    STAsum[4] <- STAsum[4] +table$STA[i]
  }
  if (Hra[5] <= table$HRA[i] && table$HRA[i] < Hra[6]){
  n[5] <- n[5] +1
  STAsum[5] <- STAsum[5] +table$STA[i]
  }
  if (Hra[6] <= table$HRA[i] && table$HRA[i] < Hra[7]){
  n[6] <- n[6] +1
  STAsum[6] <- STAsum[6] +table$STA[i]
  }
  if (Hra[7] <= table$HRA[i]){
  n[7] <- n[7] +1
  STAsum[7] <- STAsum[7] +table$STA[i]
  }
}
n  # number of people per category
STA_mean = STAsum/n
STA_mean
```


The number of people per interval is:

8 29 68 42 29 19  5

and the percentage of people who died per interval is:

0.25 0.1724 0.2059 0.1667 0.2414 0.2105 0.2

We obtain the MLE of the parameters of mean STA versus the midpoint of the HRA interval:


```{r, echo=FALSE}
mod <- glm(STA ~ HRA, family=binomial, data=table)
summary(mod)
```


The prediction equation  is:

$$
logit(\hat{\pi}) =  -1.679129 + 0.002941*Hra
$$



```{r, echo=FALSE}
table$mod <- exp(-1.679129 + 0.002941*table$HRA) / (1 + exp(-1.679129 +0.002941*table$HRA))
```


We plot  scatterplot of STA versus HRA, the mean STA versus the midpoint of the HRA interval and the estimated probabilities.


```{r, echo=FALSE}
#Hra<- c(39, 58.125,  77.25, 96.375, 115.5, 134.625, 153.75, 192)

Hra <- c( (Hra[1]+Hra[2])/2, (Hra[2]+Hra[3])/2,  (Hra[3]+Hra[4])/2, (Hra[4]+Sys[5])/2, (Sys[5]+Sys[6])/2, (Sys[6]+Sys[7])/2, (Sys[7]+Sys[8])/2, (Sys[8]+Sys[9])/2)


plot <- ggplot( data= table) + 
  geom_point(mapping = aes(x= HRA, y = STA, col = 'black')) +
      ggtitle('Vital Status over Heart Rate') +
         theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(aes(x = Hra[1], y = STA_mean[1], col = 'green')) +
  geom_point(aes(x = Hra[2], y = STA_mean[2], col = 'green')) +
  geom_point(aes(x = Hra[3], y = STA_mean[3], col = 'green')) +
  geom_point(aes(x = Hra[4], y = STA_mean[4], col = 'green')) +
  geom_point(aes(x = Hra[5], y = STA_mean[5], col = 'green')) +
  geom_point(aes(x = Hra[6], y = STA_mean[6], col = 'green')) +
  geom_point(aes(x = Hra[7], y = STA_mean[7], col = 'green')) 

plot + geom_point(aes(x=HRA, y= mod, col = 'red'), data = table) +
        scale_color_identity(breaks = c('black', 'green','red'),
                       labels = c('Response', 'Group Means', 
                                  'Estimated Probability'),
                       guide = 'legend')
```


What are the correct scales for SYS and HRA to enter the model?

The correct scale for SYS seems to be the log scale. 

HRA has a slope almost equal to zero. It seems to be that HRA is independent to STA. We will do further analysis in the following question.


## Question 6

We look at the MLE of the predictors when we include all of them:

```{r, include=FALSE}
full_mod <- glm(STA ~ AGE+SEX+RACE+SER+CAN+CRN+INF+CPR+SYS+HRA+PRE+TYP+FRA+PO2+PH+PCO+BIC+CRE+LOC, family=binomial, data=table)
```


```{r, echo=FALSE}
summary(full_mod)
```

Looking at the summary of all the predictors included, we can see that the predictors: AGE, CAN, SYS, PRE, TYP, PCO and LOC seems to be significant.



Using a model-buildigng strategy, we will select a logistic model for these predictors.



1) Let's start with the backward selection:

We begins with a complex model and sequentially removes terms. At each stage, it selects the term with the largerst p-value for removal. The process stops when any further deletion leads to a significantly poorer fit.



```{r, include=FALSE}
full_mod <- glm(STA ~ AGE+SEX+RACE+SER+CAN+CRN+INF+CPR+SYS+HRA+PRE+TYP+FRA+PO2+PH+PCO+BIC+CRE+LOC, family=binomial(link = logit), data=table)
res <- step(full_mod, scope=list(lower = ~ 1, upper = formula(full_mod)), scale = 1, trace = F, direction = "backward")
```


```{r, echo=FALSE}
res
```

2) forward selection 

At each stage, it selects the term giving the greatest improvement in fit (the term with the smallest p-value). The process stops when further additions do not significantly improve the fit.


```{r, include=FALSE}
base_mod <- glm(STA ~ 1, family=binomial, data=table)
res <- step(base_mod, scope=list(lower = ~ 1, upper =  formula(full_mod)), scale = 1, trace = F, direction = "forward")
```


```{r, echo=FALSE}
res
```


3) bestgltm with AIC



```{r, include=FALSE}
library(leaps)
library(bestglm)

#prepare the data
data <- cbind( "AGE" = table$AGE, "SEX" = table$SEX, "RACE" = table$RACE, "SER" = table$SER, "CAN" = table$CAN, "CRN" = table$CRN, "INF" = table$INF, "CPR" = table$CPR, "SYS" = table$SYS, "HRA" = table$HRA,"PRE" = table$PRE, "TYP" = table$TYP, "FRA" = table$FRA, "PO2" = table$PO2, "PH" = table$PH, "PCO" = table$PCO, "BIC" = table$BIC, "CRE" = table$CRE, "LOC" = table$LOC)

y<-table$STA
Xy <-as.data.frame(cbind(data, y))
```


```{r, echo=FALSE}
bestglm(Xy, IC = "AIC")
```



4) bestglm with BIC
```{r, echo=FALSE}
bestglm(Xy, IC = "BIC")  
```


Sum up:


1) backward selection: STA ~ AGE + CAN + SYS + TYP + PH + PCO + LOC (7 predictors)

2) forward selection: STA ~ LOC + TYP + AGE + CAN + SYS (5 predictords)

3) bestglm with AIC: STA ~ AGE + CAN + SYS + PRE + TYP + PH + PCO + LOC (8 predictors)

4) bestglm with BIC: STA ~ AGE + TYP + LOC (3 predictors)



```{r, include=FALSE}
full_mod <- glm(STA ~ AGE+SEX+RACE+SER+CAN+CRN+INF+CPR+SYS+HRA+PRE+TYP+FRA+PO2+PH+PCO+BIC+CRE+LOC, family=binomial, data=table)
back_mod <- glm(STA ~ AGE + CAN + SYS + TYP + PH + PCO + LOC, family=binomial, data=table) 
for_mod <- glm(STA ~ AGE + CAN+SYS+TYP+LOC, family=binomial, data=table)
best_AIC_mod <- glm(STA ~ AGE + CAN + SYS + PRE + TYP + PH + PCO + LOC, family=binomial, data=table)
best_BIC_mod <- glm(STA ~ AGE + TYP + LOC, family=binomial, data=table)
```

We test if the best_BIC_mod fits as well as the full model doing anova. ( we test this model first because it is composed with the fewest number of predictors with 3 predictors.)


```{r, echo=FALSE}
anova(best_BIC_mod, full_mod)
```



```{r, include=FALSE}
1 - pchisq(29.705, 17)
```

p-value(29.705, 17) = 0.0285527

Because the p - value < 0.05, we reject the null hypothesis. Therefore, we conclude that the best_BIC_mod does not fit as well as the full model.

We then test if the for_mod fits as well as the full model.


```{r, echo=FALSE}
anova(for_mod, full_mod)
```

```{r, include=FALSE}
1 - pchisq(16.273, 15)
```

p-value(16.273, 15) = 0.3641436

The p-value is > 0.05 so we do not reject the null hypothesis. Therefore, we conclude that the for_mod fits as well as the full model.

So with the forward selection, we can select the model: STA ~ AGE + CAN + SYS + TYP + LOC.

## Question 7


I -  goodness of fit 



  1) classification table

Let's evaluate the confusion matrix with a threshold = 0.5.


```{r, echo=FALSE}
threshold=0.5
predicted_values<-ifelse(predict(for_mod,type="response")>threshold,1, 0)
actual_values<-for_mod$y
conf_matrix<-table(actual_values, predicted_values)
conf_matrix
```

```{r, include=FALSE}
17/(17+23)
157/(157+3)
```

The sensitivity is 0.425 which means that the 'true positive rate' is 42.5% i.e. the probability of predicting that a patient will live (STA=0) after his admission to the ICU is .425 when the patient actually survives.

The specificity  is 0.98125 which means that the 'true negative rate' is 98% i.e. the probability of predicting that a patient will die (STA=1) after his admission to teh ICU is .98 when the patient actually dies.

With a threshold = 0.5, the specificity is better than the sensitivity.


Let's evaluate the confusion matrix with a threshold = threshold= mean(STA).


```{r, echo=FALSE}
threshold=mean(table$STA)
predicted_values<-ifelse(predict(for_mod,type="response")>threshold,1, 0)
actual_values<-for_mod$y
conf_matrix<-table(actual_values, predicted_values)
conf_matrix
```

```{r, include=FALSE}
30/(30+10)
122/(122+38)
```

The sensitivity is 0.75 which means that the 'true positive rate' is 75% i.e. the probability of predicting that a patient will live (STA=0) after his admission to the ICU is .75 when the patient actually survives.

The specificity  is 0.7625 which means that the 'true negative rate' is 76% i.e. the probability of predicting that a patient will die (STA=1) after his admission to teh ICU is .76 when the patient actually dies.

With this threshold, the sensitivity and the specificity are equivalent.



  2) ROC Curve
  
We can evaluate the model with the ROC Curve:  


```{r, include=FALSE}
library(ROCR)
rocplot=function(pred, truth, ...){
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf,...)}
```


```{r, echo=FALSE}
rocplot(predict(for_mod), for_mod$y)
```


```{r, include=FALSE}
predob = prediction(predict(for_mod), for_mod$y)
performance(predob, "auc")
```

In a Receiver Operating Characteristic (ROC) curve the true positive rate (Sensitivity) is plotted in function of the false positive rate (100-Specificity) for different cut-off points. This ROC curve is used to show in a graphical way the trade-off between sensitivity and specificity for every possible cut-off for a test or a combination of tests. The closer the curve follows the left-hand border and then the top border of the ROC space, the more accurate the test.


Concordance index = Area under curve = 0.8704687

Therefore, the proportion of pairs of data whose fitted values are pairwise concordant is higher than 87% which indicates that this model fits the data adequately and predict accordingly.


  3) Hosmer–Lemeshow test
  
```{r, include=FALSE}
library(ResourceSelection)
hl <- hoslem.test(for_mod$y, fitted(for_mod), g=10)
```


```{r, echo=FALSE}
hl
```
  
This gives p - value = 0.649 > 0.05, indicating no evidence of poor fit.
  
  
II - influential statistics  


```{r, include=FALSE}
#Influence Diagnostics 
influence.measures(for_mod)$is.inf
```

We do not observe any value where the Cook's distance and the leverage are influential. Therefore, we may not have any influential points in the data.




## Question 8

Looking at the MLE of the model we selected with forward selection, we have:

```{r, echo=FALSE}
summary(for_mod)
```


1) Find the odds ratio of survival for a 40 year old subject compared to a 50 year old subject, controlling for other factors and its 95% confidence interval. 


```{r, include=FALSE}
exp(3.832e-02*40 - 3.832e-02*50)
CI <- c(exp(-0.3832 - 1.96*1.294e-02*10), exp(-0.3832 + 1.96*1.294e-02*10))
CI
```

The odds ratio of survival (STA) for a 40 year old subject compared to a 50 year old subject, controlling for other factors is exp(3.832e-02*40 - 3.832e-02*50) = 0.6816766.

The 95% Confidence interval is [exp(-0.3832 - 1.96*1.294e-02*10), exp(-0.3832 + 1.96*1.294e-02*10)] = [0.5289698, 0.8784678].


2)  Find the odds ratio of survival and its 95% confidence interval for a subject who was in coma at ICU admission and a subject in deep stupor at admission, controlling for other factors.  

recall: LOC: Level of Consciousness 	0 = No Coma or Deep Stupor,	LOC at ICU Admission 	1 = Deep Stupor, 2= Coma

Because we are comparing LOC=2 with LOC=1, we change the baseline for LOC=2.


```{r,include=FALSE}
#define LOC as a factor with level 2 as the baseline
table$LOC <- factor(table[,21], levels=c("2", "1", "0"))
fit <- glm(table[,2]~AGE+CAN+SYS+TYP+LOC, family="binomial",data=table) 
summary(fit)
```

It turns out there is quasi complete separation for the LOC variable at level 1 (deep stupor). 

The the odds ratio of survival for a subject who was in coma at ICU admission and a subject in deep stupor at admission, controlling for other factors. is exp(1.913e+01) = 203260710.

```{r, include = FALSE}

#Profile likelihood CI
confint(fit)
```


The 95% Confidence interval is [-56.15217480, 420.351020370].
